Machine Learning for Predictive Auto-Tuning with Boosted Regression Trees

### 2011-12-14 paper/draft1.tex
m

% XXX: fix fonts in figures

@@

The empirical optimization approach, in contrast, seeks to find the best
performing code configuration by automatically generating many versions of a
parametrized kernel and benchmarking them on the actual hardware (possibly at
runtime, when contextual information about the hardware and software stack is
the richest). This method directly optimizes the metric(s) of interest (e.g. performance) and does not rely on
surrogates.  A significant advantage of such approaches is that they allow any metric to be
optimized without loss of generality. Indeed, it is possible to formulate a multi-objective optimization that minimize both run time \emph{and}
power consumption \citep{rahman2011automated}, a feat that would be even more difficult using an
analytical model-driven approach. Due to its flexibility,
empirical auto-tuning has been successfully applied to build a variety of high-performance
domain-specific libraries including dense linear algebra
\citep{clint2001automated, bilmes1997optimizing}, sparse linear algebra
\citep{vuduc2005oski}, signal processing \citep{frigo2005design}, sorting
\citep{li2004dynamically}, general stencil operations \citep{kamil2010auto}, etc.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{illustrator/fig_ntrain_munctional0_580.pdf} % on vader
\caption{The effect of the invalid configuration score ($\zeta$) and training set
size on predictive auto-tuning.
Models trained on as few as 25 configurations outperformed the reference.
Models trained on 50 or more configurations rivalled empirical auto-tuning.
Moderate values of $\zeta$ between $\log(0.25)$ and $\log(0.75)$ are best.
}
\label{fig:fig_ntrain}
\end{figure}

>>>>

@@

The empirical optimization approach, in contrast, seeks to find the best
performing code configuration by automatically generating many versions of a
parametrized kernel and benchmarking them on the actual hardware (possibly at
runtime, when contextual information about the hardware and software stack is
the richest). This method directly optimizes the metric(s) of interest (e.g. performance) and does not rely on
surrogates.  A significant advantage of such approaches is that they allow any metric to be
optimized without loss of generality. Indeed, it is possible to formulate a multi-objective optimization that minimize both run time \emph{and}
power consumption \citep{rahman2011automated}, a feat that would be even more difficult using an
analytical model-driven approach. Due to its flexibility,
empirical auto-tuning has been successfully applied to build a variety of high-performance
domain-specific libraries including dense linear algebra
\citep{clint2001automated, bilmes1997optimizing}, sparse linear algebra
\citep{vuduc2005oski}, signal processing \citep{frigo1998fftw}, sorting
\citep{li2004dynamically}, general stencil operations \citep{kamil2010auto}, etc.

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=.42]{fig_ntrain_munctional0_580.pdf} % on vader
\caption{The effect of the invalid configuration score ($\zeta$) and training set
size on predictive auto-tuning.
Models trained on as few as 25 configurations outperformed the reference.
Models trained on 50 or more configurations rivalled empirical auto-tuning.
Moderate values of $\zeta$ between $\log(0.25)$ and $\log(0.75)$ are best.
}
\label{fig:fig_ntrain}
\end{figure}
